A multi-dimensional extension of the Lightweight Temporal Compression method

### 2018-10-17 paper.tex
Some fixes

@@

 %~ is one of the linear estimation
%~ method to compress data. It is The first-order interpolator with two
%~ degrees of freedom (FOI-2DF) which is mentioned in
%~ ~\cite{jalaleddine1990ecg}.  Most of linear estimation method need a
%~ predefined parameter which is the error margin (let's call $\epsilon$)
%~ in order to guarantee the difference between estimation and original
%~ data within a scope.

%~ \todo{rewrite LTC algorithm, and there is a DP version of LTC maybe should also be written}

@@

\section{Implementation}
\label{sec:implementation}
Our code is available at ... (make a clean release!)

\todo{explain difference between exp 1 and exp 2 for euclidean}

%~ As reference above, the norm selection is importance. In this section,
%~ we show the implementation of infinity norm and Euclidean norm.

%~ \subsection{Multidimensional LTC with Manhattan distance}
%~ In section 2, we can know that the repersentation of intersection
%~ is a rectangular cuboid. So only one bounding box with n-dimensional
%~ is needed to represent the intersection. Assume a bounding box
%~ $\mathcal{BOX}_{m-1} = \bigcap_{j=1}^{m-1}{\mathcal{B}_{m-1}}$ which
%~ has Upper bound $U_i$ and lower Bound $L_i$ for each dimension
%~ $i$. So the fomular (1) in $6_{th}$ line of Algorithm 2 would
%~ be changed to $\mathcal{BOX}_{m-1} \bigcap \mathcal{B}_{m} \neq \O $. The Bounding box
%~ would be updated when this formular return true.
%~ Otherwise return false.
%~ \begin{enumerate}
    %~ \item Initialization: Let origin $(u^k_0, y^k_0) = (\tau_k, \xi_k)$.
    %~ \item For each data $(u_m^k, y_m^k)$ with diminsion $n$,
            %~ defines $(v_m, z_m) = (u_m^k - \tau^k, y_m^k - \xi^k)$ where $z_m =
            %~ \{ x_{mi} \mid 1\leqslant{i}\leqslant{n}\}$.
    %~ \item The $\mathcal{B}_{m}$ should be a ractangular cube with center
            %~ $\frac{v_1}{v_m}z_m$ and weight $\frac{v_1}{v_m}\epsilon$.
    %~ \item For each dimension $i$, box upper bound $U_i < \frac{v_1}{v_m}x_{mi}+
            %~ \frac{v_1}{v_m}\epsilon$ OR lower bound $L_i > \frac{v_1}{v_m}x_{mi}-\frac{v_1}{v_m}\epsilon$, return False.
    %~ \item Otherwise, the intersection do not be empty. Updating
            %~ the Bounding box $\mathcal{BOX}_m = \mathcal{BOX}_{m-1} \bigcap \mathcal{B}_{m}$. For
            %~ each dimension $i$ in $\mathcal{BOX}_m$, $U_i = min(U_i, \frac{v_1}{v_m}x_{mi}+
            %~ \frac{v_1}{v_m}\epsilon)$, and $L_i = max(L_i , \frac{v_1}{v_m}x_{mi}-\frac{v_1}{v_m}\epsilon)$. Return True.
%~ \end{enumerate}

%~ According the Implement, the infinity norm boils down to apply 1D LTC algorithm in each dimension respectively. Since, the intersection would be rectangle in 2D and cube in 3D, the corresponding max errors are $\frac{\epsilon}{\sqrt{2}}$ and $\frac{\epsilon}{\sqrt{3}}$. As we mentioned in Section 2, the infinity norm is more 'tolerant' than Euclidean norm. The max error between original data and reconstructed data certainly not greater than $\frac{\epsilon}{\sqrt{n}}$ where $n$ is dimension of data.

%~ \todo{maybe need a graphic}

%~ \subsection{Multidimensional LTC with Euclidean distance}
%~ In Euclidea norm, the representation of intersection is
%~ volume. There is no specific model to record repersentation
%~ with updating, so a list to save all balls $\mathcal{B}_j, 1\leqslant{j}\leqslant{m}$
%~ is needed. In order to test weather a set of balls intersect,
%~ the Helly's theorem could work, but the complexity of
%~ this method is too high to work with embeded, low-power
%~ systems which has limited CPU and RAM resource. So we
%~ implement Euclidean norm by using a method which based
%~ on plane sweep and binary search. Assume the list of balls $S_{m-1}$ contains balls from $\mathcal{B}_1$ to $\mathcal{B}_{m-1}$, and an temporary data point $output$ used for updated origin $(\tau^k, \xi^k)$. From the formula (1) in Section 3, the $(\mathcal{B}_j)_{j \in \llbracket 1, {m-1} \rrbracket}$ is a sequence of decreasing radius, so the larger balls that includes new coming ball $\mathcal{B}_{m-1}$ could be reomved from list $S_{m-1}$ so that reducing processing complexity of method. Therefore the list $S_{m-1}$ may not includes $m-1$ balls. Assume each $\Hat{\mathcal{B}} \in S$ which has corresponding center $\Hat{z}$ and radius $\Hat{\epsilon}$.
%~ Let origin $(u^k_0, y^k_0) = (\tau_k, \xi_k)$, and for each data $(u_m^k, y_m^k)$ with diminsion $n$, defines $(v_m, z_m) = (u_m^k - \tau^k, y_m^k - \xi^k)$ where $z_m = \{ x_{mi} \mid 1\leqslant{i}\leqslant{n}\}$. The $\mathcal{B}_{m}$ should be one n-ball with center $\frac{v_1}{v_j}z_m$ and radius $\frac{v_1}{v_j}\epsilon$. The method is written in Algorithm~\ref{algo:isThereIntersection}.

%~ \begin{algorithm}
    %~ \begin{flushleft}
        %~ \textbf{Input:} $S$, $\mathcal{B}_m$, $output$ \\
        %~ \textbf{Output:} $true|false$\\
    %~ \end{flushleft}
    %~ \begin{algorithmic}[1]
        %~ \State d = n \Comment{dimension index}
        %~ \State $max_d = \frac{v_1}{v_m}(x_d + \epsilon)$
        %~ \State $min_d = \frac{v_1}{v_m}(x_d - \epsilon)$
        %~ \ForAll{$\Hat{\mathcal{B}}$ \textbf{in} $S$}
            %~ \If{$\Hat{\mathcal{B}} \bigcap \mathcal{B}_m = $ \O }
                %~ \State \textbf{return} false
            %~ \ElsIf{$\mathcal{B}_m \bigcap \Hat{\mathcal{B}}  = \mathcal{B}_m $}
                %~ \State remove $\Hat{\mathcal{B}}$ from $S$
                %~ \State \textbf{continue}\\
            %~ $max_d = \min(max_d, \Hat{x_d} + \Hat{\epsilon}$)\\
            %~ $min_d = \max(min_d, \Hat{x_d} - \Hat{\epsilon}$)
            %~ \EndIf
        %~ \EndFor
        %~ \State $\textbf{add } \mathcal{B}_m \textbf{ into } S$
        %~ \If{$max_d < min_d$}
            %~ \State \textbf{return} false
        %~ \EndIf
        %~ \State temp\_output = output \Comment{used for next operation}
        %~ \If{Recursive$(min_d, max_d, n, p\_cp)$}
            %~ \State $output = temp\_output$
            %~ \State \textbf{return} true
        %~ \Else
            %~ \State \textbf{return} false
        %~ \EndIf
    %~ \end{algorithmic}
    %~ \caption{isThereIntersection}
    %~ \label{algo:isThereIntersection}
%~ \end{algorithm}

%~ In the method, selecting dimension(n as default in our method) to calculate max point value and min point value of all n-balls in this dimension. If the max value bigger than min value, then there is no intersection. After that. In Recursive function Algorithm~\ref{algo:recursive}, using binary search select
%~ median $mid_d$ of $max_d$ and $min_d$ in order to find bound range of dimension d-1 ($max_{d-1}$, $min_{d-1}$) over the plane $x_d = max_d$. If bound range of dimension $d-1$ is not empty, continue calling function \textbf{Recursive}($min_{d-1}$, $min_{d-1}$, $d-1$, $temp\_output$). Recursively calling this function, until getting result.
%~ \todo{add recursive pseudo-code}

%~ In 2-dimension Euclidean norm, the complexity of this algorithm is $O(n)+O(n\log\epsilon) = O(n\log\epsilon)$. In 3-dimension, It needs $O(\log^2\epsilon)$ to determine $mid\_3, mid\_2$, and $O(n)$ to traverse all balls in list and calculate bounding range of dimension d=1. So 3-dimensional method need $O(n \log^2\epsilon)$ totally. If we extend this idea for n-dimension, the complexiy of this method is $O(n \log^{n}\epsilon)$ where n is the number of dimension.

%~ \begin{algorithm}
    %~ \begin{flushleft}
        %~ \textbf{function} Recursive$(left, right, j, p)$  -- $j$ means $j_th$ dimension
    %~ \end{flushleft}
    %~ \begin{algorithmic}[1]
        %~ \While{$left \leqslant right$}
            %~ \State $mid \gets (left + right)/2$
            %~ \State $max \gets +\infty$
            %~ \State $min \gets -\infty$
            %~ \For{$i=1$ \textbf{to} $tmp\_list.length$}
                %~ \State $P_1$ and $P_2$ are intersection points of $tmp\_list[i]$ and line $d_{j}=mid$, $(P1.d_{j-1}\geqslant P2.d_{j-1})$
                %~ \If{$P_1.d_{j-1} < max$}
                    %~ \State $max \gets p_1.d_{j-1}$
                    %~ \State $max\_index \gets i$
                %~ \EndIf
                %~ \If{$P_2.d_{j-1} > min$}
                    %~ \State $min \gets p_2.d_{j-1}$
                    %~ \State $min\_index \gets i$
                %~ \EndIf
            %~ \EndFor
            %~ \If{$max >= min$}
                %~ \State $p.d_j \gets mid$
                %~ \If{Recursive$(min, max, j-1, p)$}
                    %~ \State \textbf{return} true
                %~ \EndIf
            %~ \EndIf
            %~ \State Assume $P_d$ is the intersection between common chord of two objects $tmp\_list[max\_index]$, $tmp\_list[min\_index]$ and their line of centers.
            %~ \If{$P_d.d_j < mid$}
                %~ \State $right \gets mid-1$
            %~ \ElsIf{$P_d.d_j > mid$}
                %~ \State $left \gets mid+1$
            %~ \EndIf

        %~ \EndWhile
        %~ \Return false
    %~ \end{algorithmic}
%~ \end{algorithm}

@@

\todo{(Put a picture of Neblina in a box worn 
on the wrist)}

@@

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-x.pdf}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-y.pdf}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-z.pdf}
\caption{Short biceps curl$^a$}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-x.pdf}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-y.pdf}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-z.pdf}
\caption{Long biceps curl}
\end{subfigure}
\caption{Time series used in Experiment 1\\
{\footnotesize $^a$ Average of az data is -7.81g. It was shifted to 0 so that the graphs can all use the same y scale.}}
\label{fig:datasets-1}
\end{figure*}

@@

\paragraph{Infinity vs Euclidean norms}
Compared to the infinity norm, the average ratio between the compression ratios obtained
with the infinity and Euclidean norm is 1.08 for 2D data, and 1.20
for 3D data. These ratios are lower than the theoretical values of
$\frac{4}{\pi}$ in 2D and $\frac{6}{\pi}$ in 3D, which are obtained for
random-uniform signals. Unsurprisingly, the infinity norm surpasses the
Euclidean norm in terms of resource consumption. Memory-wise, the
infinity norm requires a constant amount of 80~B, used to store the
intersection of n-balls. The Euclidean norm, however, uses up to 4.9~KB of memory
for the Long time series in 3D with $\epsilon$=48.8~mg. More importantly,
the amount of required memory increases for more complex time series
(Long vs Short), and it also increases with larger
values of $\epsilon$. Similar observations are made for the processing
time, with values ranging from 0.2~ms for the simplest time series and
smallest $\epsilon$, to 26.4~ms for the most complex time series and
largest $\epsilon$.Figure~\ref{fig:memory} shows the memory consumption 
of the 3D Euclidean implementation for $\epsilon$=48.8~mg: peaks appear
at the end of long compressed ranges where the signal was $\epsilon$-closed
to the linear approximation.

@@

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-x.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-y.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-z.pdf}
\caption{Walking}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-x.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-y.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-z.pdf}
\caption{Running}
\end{subfigure}
\caption{Time series used in Experiment 2}
\label{fig:datasets-2}
\end{figure*}

\begin{figure}
\includegraphics[width=\columnwidth]{./figures/memory-mohammad.pdf}
\caption{Memory usage of 3D Euclidean implementation, long biceps curl time series, $\epsilon$=48.8~mg.}
\label{fig:memory}
\end{figure}

@@

We measured the energy consumption of Neblina
for different values of $\epsilon$ and norms, to quantify the
added-value of LTC for this system.

We acquired two 3D accelerometer time series at 200Hz corresponding to
common human activities: walking and running (see Figure~\ref{fig:datasets-2}). In both cases, the
subject was wearing Neblina on their right wrist.
We collected 1,000 data points for each activity, corresponding to
5 seconds of activity.

>>>>

\newpage

@@

@@

\section{Implementation}
\label{sec:implementation}

To implement LTC with the infinity norm, we maintain a cuboid 
representation of $\cap_{l=1}^j{\mathcal{B}_l}$ throughout the 
iterations of the \texttt{while} loop in 
Algorithm~\ref{algo:general-ltc}. The implementation only requires 
constant memory and reduced CPU time to update the cuboid with the 
n-ball associated with the new data point at each iteration.

With the Euclidean norm, the intersection test is more complex, as 
mentioned previously. We keep in memory a growing set $S$ of balls and 
the bounding box $B$ of their intersection. Then, when a new point 
arrives, we consider the associated ball $\mathcal{B}_j$ and our 
intersection test works as in Algorithm~\ref{algo:euclidean}. 
\texttt{box} is a function that returns the bounding box of an n-ball. 
\texttt{find\_bisection(S, B)} searches for a point in all the elements 
in S, using plane sweep and bisection initialized by the bounds of B.

\begin{algorithm}
\begin{algorithmic}[1]
\Input
   \Desc{$S$}{Set of intersecting balls}
   \Desc{$B$}{Bounding box of the intersection of balls in $S$}
   \Desc{$\mathcal{B}_j$}{New ball to check}
\EndInput
\Output
   \Desc{$S$}{Updated set of intersecting balls}
   \Desc{$B$}{Updated bounding box}
   \Desc{$T$}{True if all the balls in S and $\mathcal{B}_j$ intersect}
\EndOutput

\If{$\exists\ \mathcal{B}_i \in S$ s.t. $\mathcal{B}_j \cap \mathcal{B}_i = \O$}
\State \Return ($S$, False)
\EndIf
\If{$\exists\ \mathcal{B}_i \in S$ s.t. $\mathcal{B}_j \subset \mathcal{B}_i$}
\State Remove $\mathcal{B}_i$ from $B$. Add $\mathcal{B}_j$ to $B$.
\State \Return ($S$, $B$, True)
\EndIf
\If{$\mathcal{B}_j \cap B = \O$}
\State \Return ($S$, $B$, False)
\EndIf
\State $B$ = box($\mathcal{B}_j$) $\ \bigcap\ $ B 
\State $S$ = $S \ \bigcup \  \{\mathcal{B}_j\}$
\State $x$ = find\_bisection($S$, $B$)
\If{$x$ == Null}
\State \Return ($S$, $B$, False)
\Else
\State \Return ($S$, $B$, True)
\EndIf
\end{algorithmic}
\caption{Intersection test for Euclidean n-balls.}
\label{algo:euclidean}
\end{algorithm}

Our code is available at 
\url{https://github.com/big-data-lab-team/stream-summarization} under
MIT license.

@@

%\todo{(Put a picture of Neblina in a box worn 
%on the wrist)}

@@

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-x.pdf}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-y.pdf}
\includegraphics[width=0.3\columnwidth]{figures/5-time-bicep-curl-plot-z.pdf}
\caption{Short biceps curl$^a$}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-x.pdf}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-y.pdf}
\includegraphics[width=0.3\columnwidth]{figures/Mohammad-bicep-curl-plot-z.pdf}
\caption{Long biceps curl}
\end{subfigure}
\caption{Time series used in Experiment 1\\
{\footnotesize $^a$ Average of az data is -7.81g. It was shifted to 0 so that the graphs can all use the same y scale.}}
\label{fig:datasets-1}
\end{figure*}

@@

\paragraph{Infinity vs Euclidean norms}
Compared to the infinity norm, the average ratio between the compression ratios obtained
with the infinity and Euclidean norm is 1.08 for 2D data, and 1.20
for 3D data. These ratios are lower than the theoretical values of
$\frac{4}{\pi}$ in 2D and $\frac{6}{\pi}$ in 3D, which are obtained for
random-uniform signals. Unsurprisingly, the infinity norm surpasses the
Euclidean norm in terms of resource consumption. Memory-wise, the
infinity norm requires a constant amount of 80~B, used to store the
intersection of n-balls. The Euclidean norm, however, uses up to 4.9~KB of memory
for the Long time series in 3D with $\epsilon$=48.8~mg. More importantly,
the amount of required memory increases for more complex time series
(Long vs Short), and it also increases with larger
values of $\epsilon$. Similar observations are made for the processing
time, with values ranging from 0.2~ms for the simplest time series and
smallest $\epsilon$, to 26.4~ms for the most complex time series and
largest $\epsilon$.
%~ Figure~\ref{fig:memory} shows the memory consumption 
%~ of the 3D Euclidean implementation for $\epsilon$=48.8~mg: peaks appear
%~ at the end of long compressed ranges where the signal was $\epsilon$-closed
%~ to the linear approximation.

@@

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{./figures/walking-1000-x.pdf}
\includegraphics[width=0.3\columnwidth]{./figures/walking-1000-y.pdf}
\includegraphics[width=0.3\columnwidth]{./figures/walking-1000-z.pdf}
\caption{Walking}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.3\columnwidth]{./figures/running-1000-x.pdf}
\includegraphics[width=0.3\columnwidth]{./figures/running-1000-y.pdf}
\includegraphics[width=0.3\columnwidth]{./figures/running-1000-z.pdf}
\caption{Running}
\end{subfigure}
\caption{Time series used in Experiment 2}
\label{fig:datasets-2}
\end{figure*}

%~ \begin{figure}
%~ \includegraphics[width=\columnwidth]{./figures/memory-mohammad.pdf}
%~ \caption{Memory usage of 3D Euclidean implementation, long biceps curl time series, $\epsilon$=48.8~mg.}
%~ \label{fig:memory}
%~ \end{figure}

@@

%~ We measured the energy consumption of Neblina
%~ for different values of $\epsilon$ and norms, to quantify the
%~ added-value of LTC for this system.

We acquired two 3D accelerometer time series at 200~Hz corresponding to
common human activities: walking and running (see Figure~\ref{fig:datasets-2}). In both cases, the
subject was wearing Neblina on their right wrist.
We collected 1,000 data points for each activity, corresponding to
5 seconds of activity.
